### 2025-04-10 chapters/2-particles.tex
Final changes to main files for first draft

\remark{}{
    The definition of fundamental particles as irreps of the Poincar\'e group actually implies that all \textit{inertial}, or non-accelerating observers, should agree on which states define a particular particle.
    %
    However, the story does not end there, and there is a forest of subtleties which lie far beyond the scope of this thesis.
    %
    One important subtlety is that in \textit{gauge theories} such as \gls{qcd} and the \gls{sm}, even the description of fundamental particles as irreducible representations of the Poincar\'e group has flaws;
    %
    in technical language, gauge invariant states of a gauge theory are associated with additional gauge fields and asymptotic charge that are not encompassed by the definition of particles as irreps of the Poincar\'e group.
}

@@

Quantum field theory (\gls{qft}) is our simplest framework for relativistic quantum mechanics, and provides elegant solutions to a huge number of conceptual difficulties.%
\footnote{
See, for example, the early chapters of Daniel Harlow's  \textit{Notes on Quantum Field Theory} \cite{}, as well as \href{https://arnold-neumaier.at/physfaq/topics/multi.html}{the blog post of Arnold Neumeier on multiparticle relativistic quantum mechanics, \texttt{available at this https url}}.
}
%
\Gls{qft} was developed first as a tool to describe the fundamental constituents of our universe -- a story that has culminated in the development of \gls{qcd} and the \gls{sm}.%
\footnote{
    And, depending on your company, string theory.
}
%
Today, however, \gls{qft} is a powerful framework for quantum mechanics even in a wide variety of observed phenomena, including non-relativistic phenomena ranging from nuclear transitions to superfluidity and superconductivity.

@@

\remark{}{
    A more in-depth exposition of \gls{qft} would require a great deal of effort, and distract from the physics of quarks, gluons, and jets that we wish to explore in this thesis.
    %
    Nonetheless, I mention that some of my favorite introductory resources on \gls{qft} are Mark Srednicki's \underline{Quantum Field Theory} \cite{Srednicki:2007qs} and Daniel Harlow's \textit{Notes on Quantum Field Theory} \cite{}.
    %
    The latter in particular gives a sense of the full, deep flavor of \gls{qft} from algebraic, calculational, and intuitive perspectives through rigorous and detailed discussion.
}

@@

Particularly notable for the case of \gls{qcd} is the beta function for the Yang-Mills coupling \(g_s\), called the \textit{strong coupling constant}.
%
It is often parameterized in terms of perturbatively calculable coefficients as
\begin{align}
    \beta_{g_s}(g_s)
    =
    -\frac{g_s}{8\pi}\,
    \sum_{n=0}^\infty \beta_n \le(\frac{g_s^2}{16\pi^2}\ri)^{n+1}
    \,,
\end{align}
\sam{Fix!!! See QCD and Collider Physics, page 25}
or
\begin{align}
    \beta_{\alpha_s}(\alpha_s)
    =
    -\sum_{n=0}^\infty \beta_n \le(\frac{\alpha_s}{4\pi}\ri)^{n+2}
\end{align}
where \(\alpha_s = g_s^2 / 4 \pi\) and the \(\beta_k\) are coefficients that can be captured by \(N^{k+1}LO\) computations.
%
This parameterization is used because the combination \(\alpha_s / 4\pi\) appears frequently in loop calculations.

In 1972-73, Gerard `t Hooft \cite{tHooft:1998qmr}, Hugh Politzer \cite{Politzer:1973fx}, and David Gross and Frank Wilczek \cite{Gross:1973ju} discovered through one-loop computations that the beta function of \gls{qcd} is negative.
%
In particular, with \(N_f\) quark flavors (i.e. up, down, charm, etc), the \glslink{accuracy}{NLO} running of the effective strong coupling constant is dictated by
\sam{Fix!!!}  % TODO: Fix beta_0, globally
\begin{align}
    \beta_0 =
    \frac{1}{4\pi}\le(\frac{11}{3} C_A - \frac{4}{3} T_F N_f\ri)
    \,,
\end{align}
with the startling result that the effective coupling
\begin{align}
    \alpha_s(\mu)
    =
    \frac{\alpha_s(\mu_0)}{1 + 2 \alpha_s(\mu_0) \beta_0 \ln\le(\mu / \mu_0\ri)}
    \,,
\end{align}
\textit{decreases as a function of the energy scale} \(\mu\).
%
This was a dramatic discovery.
%
The beta functions of known quantum field theories, such as that of the electric coupling of QED, had all been positive%
\footnote{
    To my knowledge -- with the exception of \(\phi^3\)-theory, which has the undesirable feature that its energy is unbounded below, such that it has no perturbative vacuum state.
}

@@

\remark{confinement}{
    One argument that quarks must be confined is that the physical states of our universe must be gauge invariant in the framework of gauge theories like \gls{qcd}.
    %
    Other purely theoretical hints of\gls{confinement} include \textit{renormalons} in the quark propagator \cite{}, which indicate that quarks are not physical particles in that they do not correspond to poles of the \(S\)-matrix, and the theoretical considerations of the Gribov-Zwanziger model of \gls{qcd} \cite{} that also indicate that \emph{gluons} do not correspond to poles of the \(S\)-matrix \cite{}.
    %
    Additionally, extensive studies in lattice \gls{qcd} show that \gls{confinement} can be associated with the confining \textit{Cornell potential},
    \begin{equation}
        V(r) = -\frac{\kappa}{r} + \sigma r,
    \end{equation}
    between heavy quarks, where \( r \) is the distance between the quarks, \( \kappa \) is a constant associated with the Coulomb-like part of the quark potential, and \( \sigma \) is the string tension that causes the potential to increase linearly at large distances.
    %
    This linear increase in potential leads to the \gls{confinement} of quarks, as the energy required to separate them becomes arbitrarily large.
    %
    For light quarks, the potential eventually becomes constant due to \emph{string-breaking} effects, another beautiful tale which we will not have the opportunity to discuss.
    %
    \Reff{Greensite:2011zz} discusses the linearly-confining Cornell potential in the context of \gls{qcd}, and many other aspects of\gls{confinement}.
}

@@

\remark{}{
    This precise definition, a reformulation of the definitions given in Chapter 3.1 of Weinberg's \underline{The Theory of Quantum Fields, Volume I} \cite{Weinberg:1995mt}, is taken from \href{https://www.mit.edu/~harlow/HarlowQFT1.pdf\#page=101}{Chapter 8.3 of Daniel Harlow's exceptional notes on quantum field theory}.
    %
    The basics of scattering theory in this framing discussed in depth in Prof. Harlow's notes.
    %
    I can also recommend the complementary presentations of \Reffs{Srednicki:2007qs,Schwartz:2014sze}.
}

@@

Importantly, the IR divergences of standard \glspl{qft} (including the \gls{sm} and \gls{qcd}) cancel between quantum mechanical corrections (loop integrals) and corrections due to real emissions (integrals over classical phase space) in the calculation of inclusive cross sections.
%
The proof of this result, called the \vocab{Kinoshita-Lee-Nauenberg (KLN) theorem} \cite{Kinoshita:1962ur,Lee:1964is,} (see also \Reff{Frye:2018xjj}), is \textit{far} beyond the scope of this thesis;
%
it is a giant on whose shoulders we will stand moving forward:

@@

\remark{}{
    The Fadeev-Kulish (FK) formalism for dressed, gauge invariant states \cite{}, motivated by the preliminary work of P.A.M. Dirac \cite{}, offers a complementary perspective on IR finiteness.
    %
    The FK formalism operates under the assumption that infra-red finiteness may be ensured by working with appropriately defined initial and final states involving \vocab{dressed} charged particles -- roughly, particles surrounded by clouds of gauge bosons.
    %
    See \Reffs{Frye:2018xjj,Hannesdottir:2021oqb} for a modern perspective on the FK formalism, including recent discussions of dressed states in \gls{qcd}.
}

@@

Since the standard lore is quite useful, let us present the rough argument for the idea that \gls{irc-safety} ensures perturbative computability.
%
The probability distribution for an observable \(X\),
\begin{align}
    \label{eq:prob_distribution}
    \frac{\dd\Sigma}{\dd X}(x)
    :=
    \frac{1}{\sigma}
    \int \dd \sigma
    \,\,
    \delta\le(X(\sigma) - x\ri)
    \,,
\end{align}
can be calculated in perturbation theory if and only the integral of \Eq{prob_distribution} is finite.
%
Expanding in the number of particles contributing to the cross section, we have
\begin{align}
    \label{eq:prob_distribution}
    \sigma \frac{\dd\Sigma}{\dd X}(x)
    :=
    \int \dd\sigma_2 \delta\le(X_2 - x\ri)
    +
    \int \dd\sigma_3 \delta\le(X_3 - x\ri)
    +
    \cdots
    \,,
\end{align}
where \(X_n\) is shorthand for the value of \(X\) when computed on the \(n\)-particle configuration determined by the \(n\)-particle differential cross section \(\dd \sigma_n\).

@@

The simplest form of experimentally-accessible information in particle collision events is the distribution of energy of the radiation which explodes outward from the collision;
%
we call this distribution of energy the \vocab{\gls{energy-flow}} of the collision~\cite{Sterman:1975xv, Basham:1977iq, Basham:1978bw, Basham:1978zq, Basham:1979gh,Mazzilli:2024ots,CMS:2024mlf,Tamis:2023guc}.
    %%
    %Since energy correlators can be described directly in terms of field-theoretic energy flow operators~, one can use sophisticated theoretical techniques, including the powerful technology of conformal field theories, to extract rich information about jet substructure, especially in the collinear limit~.
\begin{definitionbox}{Energy Flow}{energy-flow}
    In a particle collision with \(N\) outgoing particles with energies \(E_i\) and angles \(\hat n_i\), \(i \in \{1,\,\ldots,\,N\}\), the \gls{energy-flow} is the distribution
    %
    \begin{equation}
      \label{eq:energy-flow}
      \mc E(\hat n) = \sum_{i=1}^N E_i \delta(\hat n - \hat n_i)
      \,.
    \end{equation}

@@

\remark{}{
    If all particles are massless (and thus in the infinite-energy limit of scattering), the energy flow can be considered as a quantum operator on the Hilbert space of an underlying quantum field theory \cite{Sveshnikov:1995vi, Tkachov:1995kk, Korchemsky:1999kt, Lee:2006nr, Hofman:2008ar, Belitsky:2013xxa, Belitsky:2013bja, Kravchuk:2018htv}.
    %
    Since the energy flow depends on the energy of outgoing radiation, it may be expressed in terms of the relativistic \textit{stress tensor} \(\mathbf{\hat{\mathcal{T}}}_{\mu\nu}\), the operator which encodes energy and momentum in a relativistic \gls{qft}.
    %
    Concretely,
    \begin{align}
        \mathbf{\hat{\mathcal{E}}}(\hat{n})
        =
        \lim_{r \to \infty}
        r^2
        \int_0^\infty
        \,\,
        \dd t
        \,\,\,
        \hat{n}^i
        \,\,\,
        \mathbf{\hat{\mathcal{T}}}_{0i}
        \,
        (t,\, \vec{r} \! = \! r \hat{n})
        \,.
    \end{align}
    %
    This operator is called the \vocab{\gls{energy-flow} operator}, or sometimes the \vocab{average null energy condition (ANEC) operator}.
    %
    This operator itself plays an important role in the study of \gls{qft}.%
    \footnote{
        For example, in characterizing fundamental features of conformal field theories~\cite{Hofman:2008ar, Kologlu:2019mfz}, and even as a probe of entanglement \cite{Faulkner:2015csl,Balakrishnan:2019gxl}.
    }
    The energy flow operator may be used in particle physics applications by enabling powerful technology from conformal field theory (notably, the operator product expansion) in the computation of collider observables, especially in the collinear limit \cite{Geyer:1994zu,Dixon:2019uzg, Chen:2020vvp, Chen:2019bpb, Chen:2020adz, Chen:2021gdk, Schindler:2023cww, Gao:2023ivm, Chen:2023zlx, Chicherin:2024ifn, Chen:2024nyc, Lee:2022uwt, Chang:2022ryc,Budhraja:2024xiq,Budhraja:2024tev,Lee:2024jnt}.
    %
    The energy flow operator itself is a member of the larger class of \emph{light-ray operators} \cite{Kravchuk:2018htv,Kologlu:2019mfz}.
}

\remark{continuity}{
    A stronger requirement than \gls{irc-safety} which will also play an important role in this thesis -- particularly in \Chap{grooming} -- is that of \vocab{\gls{continuity}}.
    %
    Roughly, a collider observable \(X\) is \textit{continuous} if small changes to the energy flow of an event can only ever lead to small changes in the value of \(X\) when computed on the event.
    %
    We will use the terms \vocab{\gls{soft-discontinuity}} and \vocab{\gls{angular-discontinuity}} to refer to collider observables which exhibit discontinuous behavior when the energies or angles of particle constituents are subject to small perturbations, respectively.
    %
    We will also provide more precise definitions of continuity in the context of \gls{jet-grooming} in \Sec{emd} of \Chap{grooming}, where we discuss continuity of maps from the space of collider events to itself.
}

@@

    % Caption
    \caption[\sam{A visualization of \gls{preconfinement}}]{
        A visualization of \gls{preconfinement}, a local manifestation of \glslink{phd}{parton-hadron duality} in which partons tend to form low-mass color-singlet clusters.
        %
        \sam{complete}
    }
    \label{fig:preconfinement}
\end{figure}

@@

\Gls{preconfinement} also leads us to  the principle of \gls{lphd} \cite{Azimov:1984np,Dokshitzer:1991eq}, which roughly states that, at high enough energies, the energy flow of partons and hadrons is closely related and that even less-inclusive/\textit{local} observables have similar behavior at parton- and hadron-level.
%
Quantitative descriptions and tests of \gls{lphd} are more sparse;
%
while there is a precise OPE description of \gls{phd}, I am not aware of any such description of \gls{lphd}.
%
Nonetheless, \gls{lphd} is often phrased in terms of quantitative predictions for fragmentation functions \cite{}, particle multiplicities \cite{}, and even the by comparing parton- and hadron-/resonance-level results for locally-defined Mellin moments of structure functions \cite{}.
    % Frag?
    % https://iopscience.iop.org/article/10.1088/0954-3899/17/10/017/pdf
    % https://iopscience.iop.org/article/10.1088/0954-3899/17/10/003/pdf
    % https://link.springer.com/content/pdf/10.1007/978-1-4615-3774-8.pdf
    %
    % particle multiplicity
    % https://lib-extopc.kek.jp/preprints/PDF/1989/8902/8902234.pdf
    %
    % Local mellin
    % https://inspirehep.net/literature/781361
    % https://arxiv.org/pdf/0803.2055
and even the entanglement entropy associated with the composition of hadrons \cite{}%
\footnote{
    Thanks to Artur Avkhadiev for pointing out these references
}.
% https://arxiv.org/abs/1904.11974
% https://arxiv.org/abs/2110.06156
% https://arxiv.org/abs/2207.09430
% https://arxiv.org/abs/2305.03069
% https://arxiv.org/pdf/2408.1259
% and references therein
% https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.131.241901

@@

    \texttt{(Note:)}
    %
    To complete this exercise, I deigned to use a computer algebra system rather than my own weak hands, and the the polarization vectors and bispinors of \App{qcd-compendium} that I derived to simplify the polarization/spinor algebra of the collinear limit.
    %
    My rough code is available on \href{https://github.com/samcaf/Thesis}{\texttt{the GitHub repository associated with this thesis at this https url}}.
    %
    Spinor helicity methods can also be a useful tool, if you know them.
\end{exercise}

@@

Nonetheless, less us proceed with the \emph{formal manipulations} that will lead us to standard foundational results.
%
We now write the probability distribution to find a parton of flavor \(i\) \textit{inside} our original quark, of relative momentum fraction \(z\), as
\begin{align}
    \rho_{i \leftarrow \ell}(z)
    =
    \delta_{i\ell}
    \,
    \delta(1-z) + \sum_j \rho_{ij \leftarrow \ell}(z)
    \,
    =
    \,
    \delta(1-z)
    \,
    +
    \,
    \frac{\alpha_s}{2\pi}
    p_{j\leftarrow \ell}(z)
    \,\,
    L
    \,,
\end{align}
where we have used the definition of the \textit{inclusive} splitting function \(p\), and we have integrated the full splitting distribution.%
\footnote{
    Beyond NLO, we must include additional splitting functions with a more complicated structure.
    %
    At NNLO, for example, there are \(1 \to 3\) splitting functions describing the probability of a single parton converting into three partons \cite{}.
    %
    Nonetheless, the structure of \(\rho_{i\leftarrow \ell}\) can easily be extended, e.g. by summing over \emph{two} final-state partonic flavors in the \(1\to 3\) splitting functions to isolate the index \(i\) and, more generally, \(N-1\) final-state flavors for \(1\to N\) splitting functions.
}

@@

For gluons, our final ingredient, we can invoke symmetry under \(z \leftrightarrow 1-z\) to write \(H^{(0)}_{gg} = H^{(1)}_{gg}\).%
\footnote{
    This symmetric representation for \(p_{gg}(z)\) is non-standard, and it is common to only regularize the \(z=1\) singularity of \(p_{g\leftarrow g}(z)\).
    %
    However, we stand by the symmetry of the gluon splitting function in the physically-motivated discussion of this chapter.
    %
    We should note that using a symmetric representation of the regularized \(p_{gg}\) does not directly translate into familiar DGLAP renormalization evolution for \emph{all} structure functions.
    %
    The difference arises (\emph{only}) in the \(m\leq1\) Mellin moments (see e.g. \App{mellin}) of \(p_{gg}(z)\);
    %
    the first Mellin moment of \(p_{gg}\) in the symmetric representation we use in this thesis does not correspond exactly to an anomalous dimension of a twist-two operator associated with DGLAP evolution (see e.g. the comments of \Reff{Altarelli:1977zs} near their Eq. (37)).
    %
    However, we do not use \(m=1\) Mellin moments directly in this thesis, and the symmetric representation yields correct results for all higher-order Mellin moments.
}
%
we have no number conservation sum rule.
%
But we have \emph{the conservation of momentum}.

@@

\begin{exercise}
    Write a general form for the sum rule on splitting functions associated with a quantum number \(\chi\).
\end{exercise}

>>>>

\remark{}{
    The story does not end there, and there is a forest of subtleties which lie far beyond the scope of this thesis.
    %
    One important subtlety is that in \textit{gauge theories} such as \gls{qcd} and the \gls{sm}, even the description of fundamental particles as irreducible representations of the Poincar\'e group has flaws;
    %
    in technical language, gauge invariant states of a gauge theory are associated with additional gauge fields and asymptotic charge that are not encompassed by the definition of particles as irreps of the Poincar\'e group.
    %
    See e.g. the influential early work of \Reff{Kulish:1970ut}, which states that ``the representation of the Poincar\'e group ... acting in the physical Hilbert space ... does not contain discrete irreducible terms with nonvanishing mass. In other words, one can say that the relativistic concept of a charged particle does not exist.''
}

@@

Quantum field theory (\gls{qft}) is our simplest framework for relativistic quantum mechanics, and provides elegant solutions to a huge number of conceptual difficulties.%
\footnote{
See, for example, the early chapters of Daniel Harlow's  \textit{Notes on Quantum Field Theory} \cite{Harlow:2024zz}, as well as \href{https://arnold-neumaier.at/physfaq/topics/multi.html}{a blog post of Arnold Neumeier on multiparticle relativistic quantum mechanics, \texttt{available at this https url}}.
}
%
\Gls{qft} was developed first as a tool to describe the fundamental constituents of our universe -- a story that has culminated in the development of \gls{qcd} and the \gls{sm}.%
\footnote{
    And, depending on your company, string theory.
}
%
Today, however, \gls{qft} is a powerful framework for quantum mechanics even in a wide variety of observed phenomena, including non-relativistic phenomena ranging from nuclear transitions to superfluidity and superconductivity.

@@

\remark{}{
    A more in-depth exposition of \gls{qft} would require a great deal of effort, and distract from the physics of quarks, gluons, and jets that we wish to explore in this thesis.
    %
    Nonetheless, I mention that some of my favorite introductory resources on \gls{qft} are Mark Srednicki's \underline{Quantum Field Theory} \cite{Srednicki:2007qs} and Daniel Harlow's \textit{Notes on Quantum Field Theory} \cite{Harlow:2024zz}.
    %
    The latter in particular gives a sense of the full, deep flavor of \gls{qft} from algebraic, calculational, and intuitive perspectives through a detailed discussion that is not usually emphasized in introductory discussions of \gls{qft}.
}

@@

Particularly notable for the case of \gls{qcd} is the beta function for the Yang-Mills coupling \(g_s\), called the \textit{strong coupling constant}.
%
It is often parameterized in terms of perturbatively calculable coefficients as
\begin{align}
    \beta_{g_s}(g_s)
    =
    -g_s\,
    \sum_{n=0}^\infty \beta_n \le(\frac{g_s^2}{16\pi^2}\ri)^{n+1}
    \,,
\end{align}
or
\begin{align}
    \beta_{\alpha_s}(\alpha_s)
    =
    -2\alpha_s
    \sum_{n=0}^\infty \beta_n \le(\frac{\alpha_s}{4\pi}\ri)^{n+1}
\end{align}
where \(\alpha_s = g_s^2 / 4 \pi\) and the \(\beta_k\) are coefficients that can be captured by N\(^{k+1}\)LO computations.
%
This parameterization is used because the combination \(\alpha_s / 4\pi\) appears frequently in loop calculations.

In 1972-73, Gerard `t Hooft \cite{tHooft:1998qmr}, Hugh Politzer \cite{Politzer:1973fx}, and David Gross and Frank Wilczek \cite{Gross:1973ju} discovered through one-loop computations that the beta function of \gls{qcd} is negative.
%
In particular, with \(N_f\) quark flavors (i.e. up, down, charm, etc), the \glslink{accuracy}{NLO} running of the effective strong coupling constant is dictated by
\begin{align}
    \beta_0 = \frac{11}{3} C_A - \frac{4}{3} T_F N_f
    \,,
\end{align}
with the startling result that the effective coupling
\begin{align}
    \alpha_s(\mu)
    =
    \frac{\alpha_s(\mu_0)}{1 + \alpha_s(\mu_0) \beta_0 \ln\le(\mu / \mu_0\ri)/2\pi}
    \,,
\end{align}
\textit{decreases as a function of the energy scale} \(\mu\).
%
This was a dramatic discovery.
%
The beta functions of known quantum field theories, such as that of the electric coupling of QED, had all been positive%
\footnote{
    To my knowledge -- with the exception of \(\phi^3\)-theory, which has the undesirable feature that its energy is unbounded below, such that it has no perturbative vacuum state.
}

@@

\remark{confinement}{
    One argument that quarks must be confined is that the physical states of our universe must be gauge invariant in the framework of gauge theories like \gls{qcd}.
    %
    Other purely theoretical hints of\gls{confinement} include \textit{renormalons} in the quark propagator (see e.g. \cite{Altarelli:1995kz,Beneke:1998ui,Schindler:2023cww}), which indicate that quarks are not physical particles in that they do not correspond to poles of the \(S\)-matrix, and the theoretical considerations of the Gribov-Zwanziger model of \gls{qcd} \cite{Gribov:1977wm,Zwanziger:1988jt,Zwanziger:1989mf} that also indicate that \emph{gluons} do not correspond to poles of the \(S\)-matrix \cite{Zwanziger:1988jt} (see \Reff{Vandersickel:2012tz} for a review).
    %
    Additionally, extensive studies in lattice \gls{qcd} show that \gls{confinement} can be associated with the confining \textit{Cornell potential},
    \begin{equation}
        V(r) = -\frac{\kappa}{r} + \sigma r,
    \end{equation}
    between heavy quarks, where \( r \) is the distance between the quarks, \( \kappa \) is a constant associated with the Coulomb-like part of the quark potential, and \( \sigma \) is the string tension that causes the potential to increase linearly at large distances.
    %
    This linear increase in potential leads to the \gls{confinement} of quarks, as the energy required to separate them becomes arbitrarily large.
    %
    For light quarks, the potential eventually becomes constant due to \emph{string-breaking} effects, another beautiful tale which we will not have the opportunity to discuss.
    %
    \Reff{Greensite:2011zz} discusses the linearly-confining Cornell potential in the context of \gls{qcd}, and many other aspects of\gls{confinement}.
}

@@

\remark{}{
    This precise definition, a reformulation of the definitions given in Chapter 3.1 of Weinberg's \underline{The Theory of Quantum Fields, Volume I} \cite{Weinberg:1995mt}, is taken from \href{https://www.mit.edu/~harlow/HarlowQFT1.pdf\#page=101}{Chapter 8.3 of Daniel Harlow's exceptional notes on quantum field theory} \cite{Harlow:2024zz}.
    %
    The basics of scattering theory in this framing discussed in depth in Prof. Harlow's notes.
    %
    I can also recommend the complementary presentations of \Reffs{Srednicki:2007qs,Schwartz:2014sze}.
}

@@

Importantly, the IR divergences of standard \glspl{qft} (including the \gls{sm} and \gls{qcd}) cancel between quantum mechanical corrections (loop integrals) and corrections due to real emissions (integrals over classical phase space) in the calculation of inclusive cross sections.
%
The proof of this result, called the \vocab{Kinoshita-Lee-Nauenberg (KLN) theorem} \cite{Kinoshita:1962ur,Lee:1964is} (see also \Reff{Frye:2018xjj}), is \textit{far} beyond the scope of this thesis;
%
it is a giant on whose shoulders we will stand moving forward:

@@

\remark{}{
    The Fadeev-Kulish (FK) formalism for dressed, gauge invariant scattering \cite{Kulish:1970ut}, inspired by the preliminary work of P.A.M. Dirac \cite{Dirac:1955uv} and the later work of \Reff{Chung:1965zza}, offers a complementary perspective on IR finiteness (see also \Reff{Contopanagos:1991yb} for more developments on the KLN perspective).
    %
    The FK formalism operates under the assumption that infra-red finiteness may be ensured by working with appropriately defined initial and final states involving \vocab{dressed} charged particles -- roughly, particles surrounded by clouds of gauge bosons.
    %
    See \Reffs{Catani:1984dp,Ciafaloni:1989vs,Lavelle:1995ty,Bagan:1999jf,Larkoski:2014bxa,Frye:2018xjj,Hannesdottir:2021oqb} for additional perspectives and discussions on dressed states in \gls{qcd}.
}

@@

Since the standard lore is quite useful, let us present the rough argument for the idea that \gls{irc-safety} ensures perturbative computability.
%
The probability distribution for an observable \(X\),
\begin{align}
    \label{eq:prob_distribution_irc}
    \frac{\dd\Sigma}{\dd X}(x)
    :=
    \frac{1}{\sigma}
    \int \dd \sigma
    \,\,
    \delta\le(X(\sigma) - x\ri)
    \,,
\end{align}
can be calculated in perturbation theory if and only the integral of \Eq{prob_distribution_irc} is finite.
%
Expanding in the number of particles contributing to the cross section, we have
\begin{align}
    \sigma \frac{\dd\Sigma}{\dd X}(x)
    :=
    \int \dd\sigma_2 \delta\le(X_2 - x\ri)
    +
    \int \dd\sigma_3 \delta\le(X_3 - x\ri)
    +
    \cdots
    \,,
\end{align}
where \(X_n\) is shorthand for the value of \(X\) when computed on the \(n\)-particle configuration determined by the \(n\)-particle differential cross section \(\dd \sigma_n\).

@@

The simplest form of experimentally-accessible information in particle collision events is the distribution of energy of the radiation which explodes outward from the collision;
%
we call this distribution of energy the \vocab{\gls{energy-flow}} of the collision~\cite{Sterman:1975xv, Basham:1977iq, Basham:1978bw, Basham:1978zq, Basham:1979gh,Tamis:2023guc,Mazzilli:2024ots,CMS:2024mlf}.
\begin{definitionbox}{Energy Flow}{energy-flow}
    In a particle collision with \(N\) outgoing particles with energies \(E_i\) and angles \(\hat n_i\), \(i \in \{1,\,\ldots,\,N\}\), the \gls{energy-flow} is the distribution
    %
    \begin{equation}
      \label{eq:energy-flow}
      \mc E(\hat n) = \sum_{i=1}^N E_i \delta(\hat n - \hat n_i)
      \,.
    \end{equation}

@@

\remark{}{
    If all particles are massless (and thus in the infinite-energy limit of scattering), the energy flow can be considered as a quantum operator on the Hilbert space of an underlying quantum field theory \cite{Sveshnikov:1995vi,Tkachov:1995kk,Korchemsky:1999kt,Lee:2006nr, Hofman:2008ar, Belitsky:2013xxa, Belitsky:2013bja, Kravchuk:2018htv}.
    %
    Since the energy flow depends on the energy of outgoing radiation, it may be expressed in terms of the relativistic \textit{stress tensor} \(\mathbf{\hat{\mathcal{T}}}_{\mu\nu}\), the operator which encodes energy and momentum in a relativistic \gls{qft}.
    %
    Concretely,
    \begin{align}
        \mathbf{\hat{\mathcal{E}}}(\hat{n})
        =
        \lim_{r \to \infty}
        r^2
        \int_0^\infty
        \,\,
        \dd t
        \,\,\,
        \hat{n}^i
        \,\,\,
        \mathbf{\hat{\mathcal{T}}}_{0i}
        \,
        (t,\, \vec{r} \! = \! r \hat{n})
        \,.
    \end{align}
    %
    This operator is called the \vocab{\gls{energy-flow} operator}, or sometimes the \vocab{average null energy condition (ANEC) operator}.
    %
    This operator itself plays an important role in the study of \gls{qft}.%
    \footnote{
        For example, in characterizing fundamental features of conformal field theories~\cite{Hofman:2008ar, Kologlu:2019mfz}, and even as a probe of entanglement \cite{Faulkner:2015csl,Balakrishnan:2019gxl}.
    }
    The energy flow operator may be used in particle physics applications by enabling powerful technology from conformal field theory (notably, the operator product expansion) in the computation of collider observables, especially in the collinear limit \cite{Geyer:1994zu,Dixon:2019uzg, Chen:2020vvp,Chen:2019bpb, Chen:2020adz,Chen:2021gdk, Schindler:2023cww, Gao:2023ivm, Chen:2023zlx, Chicherin:2024ifn, Chen:2024nyc, Lee:2022uwt,Chang:2022ryc,Budhraja:2024xiq,Budhraja:2024tev,Lee:2024jnt}.
    %
    The energy flow operator itself is a member of the larger class of \emph{light-ray operators} \cite{Kravchuk:2018htv,Kologlu:2019mfz}.
}

\remark{continuity}{
    A stronger requirement than \gls{irc-safety} which will also play an important role in this thesis -- particularly in \Chap{grooming} -- is that of \vocab{\gls{continuity}}.
    %
    Roughly, a collider observable \(X\) is \textit{continuous} if small changes to the energy flow of an event can only ever lead to small changes in the value of \(X\) when computed on the event.
    %
    We also use the terms \vocab{\gls{soft-discontinuity}} and \vocab{\gls{angular-discontinuity}} to refer to collider observables which exhibit discontinuous behavior when the energies or angles of particle constituents are subject to small perturbations, respectively.
    %
    These concepts play a particularly central role in \Chap{grooming}, where we also provide more precise definitions of continuity in the context of \gls{jet-grooming} in \Sec{emd}, with a focus on the continuity of maps from the space of collider events to itself (i.e. maps \(f:S \to S\) in the language of \Def{observable}).
}

@@

    % Caption
    \caption[A visualization of \gls{preconfinement}\sam{incomplete}]{
        A visualization of \gls{preconfinement}, a local manifestation of \glslink{phd}{parton-hadron duality} in which partons tend to form low-mass color-singlet clusters. \sam{complete}
    }
    \label{fig:preconfinement}
\end{figure}

@@

\Gls{preconfinement} also leads us to  the principle of \gls{lphd} \cite{Azimov:1984np,Dokshitzer:1991eq}, which roughly states that, at high enough energies, the energy flow of partons and hadrons is closely related and that even less-inclusive/\textit{local} observables have similar behavior at parton- and hadron-level.
%
Quantitative descriptions and tests of \gls{lphd} are more sparse;
%
while there is a precise OPE description of \gls{phd}, I am not aware of any such description of \gls{lphd}.
%
Nonetheless, \gls{lphd} is often phrased in terms of quantitative predictions for fragmentation functions \cite{Dokshitzer:1991eq,Dokshitzer:1991fc}, particle multiplicities \cite{Andersson:1989ww}, or more recently by comparing parton- and hadron-/resonance-level results for locally-defined Mellin moments of structure functions \cite{Psaker:2008ju}.
%
\gls{lphd} is also a useful starting point for discussing hadrons as entangled states of quarks and gluons, and even for extracting information about the inner entanglement of hadrons in terms of collision measurements \cite{Tu:2019ouv,Hentschinski:2021aux,Hentschinski:2022rsa,Hentschinski:2023izh,Hentschinski:2024gaa}.%
\footnote{
    Thanks to Artur Avkhadiev for pointing out this feature of \gls{lphd}.
}.

@@

    \texttt{(Note:)}
    %
    To complete this exercise, I deigned to use a computer algebra system rather than my own weak hands, and the the polarization vectors and bispinors of \App{qcd-compendium} that I derived to simplify the polarization/spinor algebra of the collinear limit.
    %
    My rough code (which \textit{does not} involve importing more dedicated technology, like FeynCalc, etc.) is available on \href{https://github.com/samcaf/Thesis}{\texttt{the GitHub repository associated with this thesis at this https url}}.
    %
    Spinor helicity methods can also be a useful tool, if you know them.
\end{exercise}

@@

Nonetheless, less us proceed with the \emph{formal manipulations} that will lead us to standard foundational results.
%
We now write the probability distribution to find a parton of flavor \(i\) \textit{inside} our original quark, of relative momentum fraction \(z\), as
\begin{align}
    \rho_{i \leftarrow \ell}(z)
    =
    \delta_{i\ell}
    \,
    \delta(1-z) + \sum_j \rho_{ij \leftarrow \ell}(z)
    \,
    =
    \,
    \delta(1-z)
    \,
    +
    \,
    \frac{\alpha_s}{2\pi}
    p_{j\leftarrow \ell}(z)
    \,\,
    L
    \,,
\end{align}
where we have used the definition of the \textit{inclusive} splitting function \(p\), and we have integrated the full splitting distribution.%
\footnote{
    Beyond NLO, we must include additional splitting functions with a more complicated structure.
    %
    At NNLO, for example, there are \(1 \to 3\) splitting functions describing the probability of a single parton converting into three partons \cite{Campbell:1997hg,Catani:1999ss} (see also \Reff{Larkoski:2024uoc} and references therein).
    %
    Nonetheless, the structure of \(\rho_{i\leftarrow \ell}\) can easily be extended, e.g. by summing over \emph{two} final-state partonic flavors in the \(1\to 3\) splitting functions to isolate the index \(i\) and, more generally, \(N-1\) final-state flavors for \(1\to N\) splitting functions.
}

@@

For gluons, our final ingredient, we can invoke symmetry under \(z \leftrightarrow 1-z\) to write \(H^{(0)}_{gg} = H^{(1)}_{gg}\).%
\footnote{
    This symmetric representation for \(p_{gg}(z)\) is non-standard, and it is common to only regularize the \(z=1\) singularity of \(p_{g\leftarrow g}(z)\).
    %
    However, we stand by the symmetry of the gluon splitting function in the physically-motivated discussion of this chapter.
    %
    We should note that, according to \Reff{Altarelli:1977zs}, using a symmetric representation of the regularized \(p_{gg}\) does not directly translate into familiar DGLAP renormalization evolution for \emph{all} structure functions (see discussion near their Eq. (37)).
    %
    However, differences due to our symmetric representation here arise (\emph{only}) in the \(m\leq1\) Mellin moments (see e.g. \App{mellin}) of \(p_{gg}(z)\).
    %
    We do not use \(m=1\) Mellin moments directly in this thesis, and the symmetric representation yields correct results for all higher-order Mellin moments.
    %
    Furthermore, since the Mellin transform of the symmetric and standard representations, the \(\hat{p}_{gg}(m)\), are analytic and agree for \(\Re m > 1\), the identity theorem suggests that they must have the same analytic continuation.
}
%
we have no number conservation sum rule.
%
But we have \emph{the conservation of momentum}.

@@

\begin{exercise}
    Write a general form for the sum rule on splitting functions associated with a conserved quantum number \(\chi\).
\end{exercise}
